---
title: "Exercises Assignment #6: HOMEWORK"
author: "Julia Kwiecinska"
format: html
editor: visual
---

# Libraries

Here I load libraries I will use in this file.

```{r}
library(rmarkdown)
library(tidyr)
library(dplyr)
```

# Data

For this homework I will be working on **iris** dataset.

## Load iris dataset

Here I load a dataset called "iris" with measurements on 3 iris species.

```{r echo=TRUE}
data("iris")

paged_table(iris)
```

## Transform it to longer

Here I transform iris dataset to longer by using tidyr::pivot_longer() to get 3 columns - Species \| Trait \| Size.

```{r}
iris_long <- iris %>% 
  tidyr::pivot_longer(cols=matches("Sepal|Petal"), names_to = "Trait", values_to = "Size") 

iris_long
```

# Questions!

## What is the mean length and width of the sepal and petal of each species?

After grouping by Species and Trait I calculate the mean for each trait (Petal.Length, Petal.Width, Sepal.Length, Sepal.Width) for each Species.

```{r echo=TRUE}
iris_long1 <- iris_long %>%
  group_by(Species, Trait) %>% 
  summarise(MeanIRIS = mean(Size)) 

iris_long1
```

## What is the largest and widest, petal and sepal, of virginica and versicolor species?

Here I filter two of the species by using dplyr::filter(). After that I group by Species and Trait as before and summarise looking for max size in each of traits.

```{r echo=TRUE}
iris_long2 <- iris_long %>% 
  dplyr::filter(Species %in% c("virginica", "versicolor")) %>% 
  group_by(Species, Trait) %>% 
  summarise(MaxSize = max(Size)) 

iris_long2
```

## For each species, for sepal and petal, width and length, how many observations are above the mean?

Here I'm using left_join() to join the mean values to the original long dataset based on species and trait. Then I count the observations above the mean for each species, for sepal and petal width and length.

```{r echo=TRUE}

iris_above_mean <- iris_long %>%
  left_join(iris_long1, by = c("Species", "Trait")) %>%
  group_by(Species, Trait) %>%
  summarise(obs_above_mean = sum(Size > MeanIRIS))

iris_above_mean

```
